---
title: "code_for_final"
author: Ethan Blacher, Maddie Manzagol, Mary Cadogan
format: 
  html:
    toc: true
    toc-location: left
    code-fold: true
    theme: yeti 
execute:
  message: false
  warning: false
---

Problem 1

```{r Install_Packages}
# should haves (from last week)
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) # or equivalent
library(flextable) # or equivalent
library(car)
library(broom)
# would be nice to have
library(corrplot)
library(AICcmodavg)
library(GGally)
library(MuMIn)
```

Problem 1: How does total number differ between kangaroo rat mound locations?

Introduction

Methods

Results

```{r Read_Data}
kangaroo <- read_csv(here("data", "prob1_data.csv"))%>% 
  #make the column names clear
  clean_names() %>%
    select(loc, seeds)
```

```{r}
gg_miss_var(kangaroo)
```

Null: Kangaroo mound locations has no impact on total seed numbers

Alternate: Kangaroo mound locations has an impact on total seed numbers



Problem 2: How does seed count vary with plot type (shrub or open), plant species, and total number of inflorescences? Is there a simpler model that explains seed count, and if so, what is it?

Introduction




```{r Read_Data}
flower <- read_csv(here("data", "prob2", "shrubstudy_seed_ctwt.ms.data.csv"))%>%
    #make the column names clear
  clean_names()%>%
  select(treatment, species, total_nr_infl)
```

```{r}
#check for missing variables
gg_miss_var(flower)
```

```{r}

```

