---
title: "code_for_final"
author: Ethan Blacher, Maddie Manzagol, Mary Cadogan
format: 
  html:
    toc: true
    toc-location: left
    code-fold: true
    theme: yeti 
execute:
  message: false
  warning: false
---

Problem 1

```{r Install_Packages}
# should haves (from last week)
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) # or equivalent
library(flextable) # or equivalent
library(car)
library(broom)
# would be nice to have
library(corrplot)
library(AICcmodavg)
library(GGally)
library(MuMIn)
```

<<<<<<< HEAD
Problem 1: How does total number differ between kangaroo rat mound locations?

Introduction
=======
Introduction:  
Write a 5-10 sentence introduction to the analysis you conduct. Some useful information
could include (but is not limited to):
a. What are Kangaroo rats?
b. Why might people care about them?
c. Why do you think it could be useful and/or important to predict individual biomass
from morphological, physiological, or taxonomic characteristics?
d. What questions are you addressing?
e. What hypotheses are you testing?
f. Include 3-5 in-text citations of peer-reviewed work.

In order to ensure plant species diversity within a given environment, the persistence of a diverse seed bank is essential as oftentimes seed bank diveristy is a key indicator to population persistence (J.P. Bakker). With that said, seed survival, which is largely dictated by a diverse seed bank, is highly correlated with population growth rates for many plant species (Adams and Marsh). One way that seed survival and, consequently, seed bank diversity is encouraged is through disturbance, as small-scale disturbances to an ecosystem are known to increase heterogeneity and plant diversity within that ecosystem (Koontz and Simpson). In the Chihuahuan Desert grassland, kangaroo rat mounds are considered small-scale disturbances as these mounds provide micro-habitats that create "safe-spaces" for seed accumulation where plant populations can establish, increasing plant diversity and community structure (Koontz and Simpson). With that being said, this study aims to uncover how total seed count differs between different kangaroo rat mound locations to gain a better understanding of what specific properties of kangaroo rat mounds aid in diversifying seed plants and adult plant communities. 

this will likely need some work but just a start^

>>>>>>> 2d14c2fab891db0fffea6837b2d23e9d05966ab9

Methods

Results

```{r Read_Data}
kangaroo <- read_csv(here("data", "prob1_data.csv"))%>% 
  #make the column names clear
  clean_names() %>%
    select(loc, seeds)
```

```{r}
gg_miss_var(kangaroo)
```

Null: Kangaroo mound locations has no impact on total seed numbers

Alternate: Kangaroo mound locations has an impact on total seed numbers



Problem 2: How does seed count vary with plot type (shrub or open), plant species, and total number of inflorescences? Is there a simpler model that explains seed count, and if so, what is it?

Introduction




```{r Read_Data}
flower <- read_csv(here("data", "prob2", "shrubstudy_seed_ctwt.ms.data.csv"))#%>%
    #make the column names clear
#  clean_names()%>%
#  select(treatment, species, total_nr_infl) mabye add some other columns
```

```{r}
gg_miss_var(flower)
```

```{r}
#check for missing variables
gg_miss_var(flower)
```

```{r}

```

