---
title: "code_for_final"
author: Ethan Blacher, Maddie Manzagol, Mary Cadogan
format: 
  html:
    toc: true
    toc-location: left
    code-fold: true
    theme: yeti 
execute:
  message: false
  warning: false
---

Problem 1

```{r Install_Packages}
# should haves (from last week)
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) # or equivalent
library(flextable) # or equivalent
library(car)
library(broom)
# would be nice to have
library(corrplot)
library(AICcmodavg)
library(GGally)
library(MuMIn)
```

Introduction:\
Write a 5-10 sentence introduction to the analysis you conduct. Some useful information could include (but is not limited to): a. What are Kangaroo rats? b. Why might people care about them? c. Why do you think it could be useful and/or important to predict individual biomass from morphological, physiological, or taxonomic characteristics? d. What questions are you addressing? e. What hypotheses are you testing? f. Include 3-5 in-text citations of peer-reviewed work.

In order to ensure plant species diversity within a given environment, the persistence of a diverse seed bank is essential as oftentimes seed bank diveristy is a key indicator to population persistence (J.P. Bakker). With that said, seed survival, which is largely dictated by a diverse seed bank, is highly correlated with population growth rates for many plant species (Adams and Marsh). One way that seed survival and, consequently, seed bank diversity is encouraged is through disturbance, as small-scale disturbances to an ecosystem are known to increase heterogeneity and plant diversity within that ecosystem (Koontz and Simpson). In the Chihuahuan Desert grassland, kangaroo rat mounds are considered small-scale disturbances as these mounds provide micro-habitats that create "safe-spaces" for seed accumulation where plant populations can establish, increasing plant diversity and community structure (Koontz and Simpson). With that being said, this study aims to uncover how total seed count differs between different kangaroo rat mound locations to gain a better understanding of what specific properties of kangaroo rat mounds aid in diversifying seed plants and adult plant communities.

this will likely need some work but just a start\^

Methods

Results

```{r Read_Data}
# Read the CSV file into a data frame
kangaroo <- read_csv(here("data", "prob1_data.csv"))
# Define values to remove from the 'loc' column
values_to_remove <- c("I", "E")
# Remove rows with specific values in the 'loc' column
kangaroo_filtered <- kangaroo[!(kangaroo$loc %in% values_to_remove), ]
# Define values to remove from the 'species' column
other_values_to_remove <- c("soil", "dist", "gravel", "litter", "plant")
# Remove rows with specific values in the 'species' column
kangaroo_filtered <- kangaroo_filtered[!(kangaroo_filtered$species %in% other_values_to_remove), ]
# Calculate the sum of 'seeds' column based on different values in the 'mnd' column
subtotal <- aggregate(seeds ~ mnd, data = kangaroo_filtered, FUN = sum)
```

```{r}
#check for missing data
gg_miss_var(kangaroo)
```

Assumptions: 

We used the Kruskal-Wallis test because it is a non-parametric alternative to the one-way ANOVA that accounts for discrete data. The test tells us if there are significant differences among mounds, but does not reveal which the differences between individual mounds. 
The data meets the assumptions for the Kruskal-Wallis test, because it has the categorical predictor variables (mound locations), the samples were collected independently, and each mound has at least 5 observations.




Null: Kangaroo mound locations has no impact on total seed numbers

Alternate: Kangaroo mound locations has an impact on total seed numbers

Problem 2.

```{r Read_Data}
flower <- read_csv(here("data", "prob2", "shrubstudy_seed_ctwt.ms.data.csv"))#%>%
    #make the column names clear
#  clean_names()%>%
#  select(treatment, species, total_nr_infl) mabye add some other columns
```

```{r}
gg_miss_var(flower)
```
