---
title: "FINAL"
author: Ethan Blacher, Maddie Manzagol, Mary Cadogan
format: 
  html:
    toc: true
    toc-location: left
    code-fold: true
    theme: yeti 
execute:
  message: false
  warning: false
---

# Problem 1

How does total seed number differ between kangaroo rat mound locations?

## Introduction

In order to ensure plant species diversity within a given environment, the persistence of a diverse seed bank is essential as oftentimes seed bank diveristy is a key indicator to population persistence (Bakker et al., 1996), with the growth rates of many plant species being largely dictated by seed survival (Adams and Marsh, 2005). One way that seed survival and, consequently, seed bank diversity is encouraged is through disturbance, as small-scale disturbances to an ecosystem are known to increase heterogeneity and plant diversity within that ecosystem (Koontz and Simpson, 2010). In the Chihuahuan Desert grassland, kangaroo rat mounds are considered small-scale disturbances as these mounds provide micro-habitats that create "safe-spaces" for seed accumulation where plant populations can establish, increasing plant diversity and community structure (Koontz and Simpson, 2010). With that being said, this study aims to uncover how total seed count differs between different kangaroo rat mound locations (base, surrounding, edge, and interspace)to gain a better understanding of what specific properties of kangaroo rat mounds aid in diversifying seed plants and adult plant communities. We are testing the null hypothesis that total seed count does not vary by mound location against the alternative hypothesis that there is a statistically significant difference in seed counts between mound locations.

## Methods


Step for cleaning/wrangling: 

In order to organize the data after reading in the necessary packages, it is important to see what data was measured. To do so, the respective csv file was read in using the read_csv and the here function and saved as an object titled kangaroo. We then further cleaned the data set by mutating it to get rid of direction, nameing this data set  "kangaroo_summary". Then we cleaned again creating a thrid data set titled "kangaroo_summary_2", which removed the percent count data in the species column (this included "soil", "dist", "gravel", "litter", "plant") and renamed column names and objects within the columns to be more clear for readability as well as for graph formating. Lastly we created a final data set titled "subtotal" which aggregated the total seed count per location so that this could be visually displayed in a graph further down in the analysis. 

```{r}
# load in packages
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) 
library(flextable)
library(car)
library(broom)
library(corrplot)
library(AICcmodavg)
library(GGally)
library(MuMIn)
library(MASS) 
library(lme4)
library(glmmTMB) 
library(DHARMa)
library(lmtest)
library(skimr)
```

```{r}
kangaroo <- read_csv(here("data", "prob1_data.csv")) 
# Read the CSV file into a data frame
kangaroo_summary <- kangaroo %>% 
  #get rid of dir column
  mutate(dir = NULL) 
#create object of all rows containing these values from the species column, their results were presented as percentages and messed up data
other_values_to_remove <- c("soil", "dist", "gravel", "litter", "plant")

#remove the values listed above to create final object 
kangaroo_summary2 <- kangaroo_summary[!kangaroo_summary$species %in% other_values_to_remove, ] %>% 
#rename items that will be used later in graphs
  rename(Location = loc, Mounds = mnd) %>% 
  mutate(Location = case_when(
    Location == "B" ~ "Base", 
    Location == "D" ~ "Surrounding",
    Location == "E" ~ "Edge", 
    Location == "I" ~ "Interspace"))

#create an object that has a total seed count per location 
subtotal <- aggregate(seeds ~ Location, data = kangaroo_summary2, FUN = sum)
  
```

**Figure 1: Visualization for missing variables**

We ran the gg_miss_var() function in order to check for missing data; however, after running the function we saw no missing data. 
```{r}
#check for missing data
gg_miss_var(kangaroo_summary2)
```

**Figure 2: Visual checks for normal distribution of data** Points in the Quartile-Quartile Plot (QQ Plot) represent sample quantiles compared against theoretical quantiles from a normal distribution. The red line represents a 1:1 relationship between sample and theoretical quantiles.

```{r}
#create qq plot to get the distribution of data for each location
ggplot(data= kangaroo_summary2, aes(sample = seeds)) +
  stat_qq()+
  stat_qq_line(aes(sample = seeds), color = "red")+
  facet_wrap(~Location) +
  labs(title = "Distribution of Data Per Mound Location", y = "Seed Count")
```

**Table 1: Levene Test** The following table assesses the equality of variances across the four different mound locations.

```{r}
#run a levene test using leveneTest to check variances
levene_test <- leveneTest(seeds ~ Location, data = kangaroo_summary2)

#turn results into readable flextable 
levene_table <- tidy(levene_test) %>% 
#clean up table
  mutate(statistic = round(statistic, digits = 3)) %>% 
  mutate(p.value = case_when(p.value < 0.001 ~ "< 0.001")) %>% 
  flextable() %>% 
  #change header labels
  set_header_labels(statistic = "F-statistic", 
                    df.residual = "Residual Degrees of Freedom", 
                    df = "Degrees of freedom",
                    p.value = "p-value") %>%
  autofit()

levene_table
```

To address the question if the four different kangaroo rat mound locations have even total seed count variances across, we applied a Levene’s test, which assumes independent observations as well as quantitative data. Levene's test showed that the variances for total seed count at different kangaroo rat mound locations were not equal, F(3) = 14.936, p = <0.001.



We used the Kruskal-Wallis test because it is a non-parametric alternative to the one-way ANOVA that accounts for discrete data. This test tells us if there are significant differences between total seed count across different mound locations; however, it does not reveal what the differences in total seed count are between different mound location. Our data meets the assumptions for the Kruskal-Wallis test, because the predictor variables (mound locations) is categorical, the samples were collected independently, and each mound location has over 5 observations. Since the Kruskal test only tells us whether there is a significant difference in total seed count between different mound locations, we also performed a post hoc test a multiple comparison that highlights which groups, mound locations,  differ from each other.

**Table 2: Kruskal-Wallist Test results** to determine the difference in ranks between the four locations.

```{r}
#run kruskal.test to determine if there is a signficant difference among the 4 locations
kruskal_test <- kruskal.test(seeds ~ Location, data = kangaroo_summary2)

#create readable flextable
kruskal_table <- tidy(kruskal_test) %>% 
  mutate(statistic = round(statistic, digits = 3)) %>% 
  mutate(p.value = round(p.value, digits = 4)) %>% 
  flextable() %>% 
  #change header labels
    set_header_labels(statistic = "H-statistic", 
                    parameter = "Parameter", 
                    method = "Method",
                    p.value = "p-value") %>%
  autofit()

kruskal_table
summary(kruskal_test)
```

**Table 3: Pairwise Wilcox Test results** to further analyze signficant differences between all combinations of locations. A post-hoc test to the Kruskal-Wallis test.

```{r}
#run post hoc test using pairwise.wilcox.test to compare between all combinations of locations, add bonferroni corrections
wilcox_test <- pairwise.wilcox.test(kangaroo_summary2$seeds, kangaroo_summary2$Location, p.adjust.method = "bonferroni")

#create readable flextable
wilcox_table <- tidy(wilcox_test) %>% 
#clean up table
  mutate(p.value = round(p.value, digits = 4)) %>% 
  flextable() %>% 
  #change headerl labels
    set_header_labels(group1 = "Group 1", 
                      group2 = "Group 2",
                      p.value = "p-value") %>%
  autofit()

wilcox_table
summary(wilcox_test)
```

**Table 4: Negative Binomial Regression Model** to test the significance of estimated coefficients between the four locations. The base location is set as the Intercept, with the estimate coefficients representing the strength and direction of relationship (how much a one unit increase in predictor variable has on the outcome variables).

```{r}
#create negative binomial model using glm.nb 
negbinomial_model <- glm.nb(seeds ~ Location, data = kangaroo_summary2)

#create flextable for negative bonimial results 
negbinomial_table <- tidy(negbinomial_model) %>% 
#clean up table
  mutate(p.value = round(p.value, digits = 5)) %>% 
  mutate(estimate = round(estimate, digits = 3)) %>% 
  mutate(std.error = round(std.error, digits = 3)) %>% 
  mutate(statistic = round(statistic, digits = 3)) %>% 
  flextable() %>% 
  #change header labels
  set_header_labels(statistic = "T-statistic", 
                    estimate = "Estimate Coefficient", 
                    term = "Location", 
                    std.error = "Standard Error",
                    p.value = "p-value") %>%
  autofit()

negbinomial_table 
```

```{r}
#look at details of negative binomial model
summary(negbinomial_model)

#getting IRR values using the exp(coef()) test
exp_coef <- exp(coef(negbinomial_model))

exp_coef
```

## Results

An analysis using the Kruskal-Wallis test was conducted to assess whether there was a significant difference in total seed count between different locations on the mound. We tested the null hypothesis that different mound locations did not predict a significant difference in total seed count, while the alternative hypothesis suggested that different mound locations did predict a significant difference in total seed count. The Kruskal-Wallis test proved that there is a significant difference in total seed count between different mound locations (H(3)= 14.438, p<0.001). After conducting the Kruskal-Wallis test and finding a significant difference in the total seed count among different mound locations, post hoc analysis using a Pairwise Wilcox Test was performed to determine the specific differences in total seed count among the different mound locations. The following differences were observed: there was no significant difference in total seed count between mound locations B and D (p = 1), there was no significant difference in total seed count between mound locations E and B (p=0.808), there was no significant difference in total seed count between mound locations E and D (p=1), there was a significant difference in total seed count between mound locations I and B (p= 0.0018), there was a significant difference in total seed count between mound locations I and D (p=0.054), and lastly there was no significant difference between total seed count between mound locations I and E (p= 0.083). Compared to location B (base), location D (surround)  is associated with a 0.938 decrease in the log of the total seed count. This corresponds to a 0.3913 increase in the expected total seed count at the surrounding location of the mound for every 1 unit of increase in total seed count for the base of the mound. Compared to the base, the edge (location E) is associated with a 1.884 decrease in the log of the total seed count. This corresponds to a 0.1519 increase in the expected total seed count at the surrounding location of the mound for every 1 unit of increase in total seed count for the base of the mound. Compared to the base, the interspace (location I) is associated with a 2.872 decrease in the expected total seed count. This corresponds to a 0.0565 increase in the expected total seed count at the interspace location of the mound for every 1 unit of increase in total seed count for the base of the mound. Given that the respective p-values for each comparison were less than 0.05, we can consider the mound location significantly predicts the total seed count (p < 0.001 for all locations comparatively to the base). All in all, the statistical test we ran highlighted that the location on the mound can impact the total seed count, indicating that small-scale disturbances do impact the volume of a seed bank and its diveristy. 

**Figure 3: Differences in total seed count at each mound location seperated by all the individual mounds that were sampled.** Each plot is representative of the different mounds that were sampled and are presented by their associated plot numbers.

```{r}
#create column plot to compare across the different mounds 
ggplot(kangaroo_summary2, aes(x = Location, y = seeds, fill = Location)) +
  geom_col() +
#change fill colors of locations
  scale_fill_manual(values = c("lightblue", "pink", "lavender", "grey")) + 
  #add labels
  labs(y = "Seeed Counts ", title = "Seed Abundance Per Location Separated by the Different Mounds", caption = "Data source: Koontz and Simpson (2010), The composition of seed banks on kangaroo rat (Dipodomys spectabilis) mounds in a Chihuahuan Desert grassland") + 
  #separate by the different mounds
  facet_wrap(~Mounds) +
  theme_classic() +
  theme(#add and edit legend 
        legend.position = c(0.85, 0.13),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 12),
        #get rid of grid marks
        panel.grid = element_blank(), 
        #get rid of axis ticks and text
        axis.ticks.x = element_blank(),  
        axis.text.x = element_blank(),
        #change font, position, and sizes of axises, caption, and title
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 14, hjust = 0.5),
        plot.caption = element_text(size = 8, hjust = 0.5, face = "italic"), 
        plot.title.position = "plot",
        text = element_text(family = "Times New Roman"))
  
```

**Figure 4: Difference in the sum total seed count at each mound location across all sampled mounds.**

```{r}
#create column plot of subtotal data
ggplot(subtotal, aes(x = Location, y = seeds, fill = Location)) +
  geom_col() +
  #add labels
  labs(x = "Mound Locations", y = "Seed Count", caption = "Data source: Koontz and Simpson (2010), The composition of seed banks on kangaroo rat (Dipodomys spectabilis) mounds in a Chihuahuan Desert grassland", title = "Seed Abundance By Varying Mound Locations", color = "Location") +
  theme_classic() +
  #change fill colors of locations
  scale_fill_manual(values = c("lightblue", "pink", "lavender", "grey")) +
  theme(#create and edit legend
        legend.position = c(0.85, 0.7),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        #get rid of grids 
        panel.grid = element_blank(), 
        #change font, position, and sizes of axises, caption, and title
        axis.text = element_text(size = 14), 
        axis.title = element_text(size = 16),
        plot.title = element_text(size = 18, hjust = 0.5),
        plot.caption = element_text(size = 12, hjust = 0.5, face = "italic"), 
        plot.title.position = "plot",
        text = element_text(family = "Times New Roman"))
```



## Problem 2

How does seed count vary with plot type (shrub or open), plant species, and total number of inflorescences? Is there a simpler model that explains seed count, and if so, what is it?

## Introduction

Determining the major contributors to higher seed abundance and diversity levels is crucial for maintaining and expanding the growth of native plant communities. This study took place in Newot Ridge, Colorado where shrubification has been increasing over the past 50 years due to warming and nitrogen desposition, thus changing the microclimates within the alpine ecosytem (Seaver, 2020) by altering species richness and diversity (Zehnder et al., 2020), providing more coverage for local fauna, and covering the ground with litter (Seaver, 2020). Additionally, shrub presence has been found to impact seed dispersal through a variety of mechanisms, such as wind protection gravity rolling, herbivory, and ground cover, that have the potential to etiher increase or decrease seed abundance depending on the observed area (Giladi et al., 2013). Lastly, seed yields have been found to be heavily dependent on the total number of inflorescences in a flowering period (Li et al., 2023), with inflorescences typically varying by plant species as well. This study seeks to determine how plot type (shrub or open), plant species, and total number of inflorescences impacts seed abundance, with the null hypothesis being that seed abundance is independent from all the proposed predictor variables.

## Methods

In order to organize the data after reading in the necessary packages, it is important to see what data was measured. To do so, the seed count csv file was read in using the read_csv and the here function. We then made two different data sets, flower_viz to see missing data, flower, which cleaned the names and dropped the missing data for nr_seeds which was pertinent to our model and visualization and flower_subset, which dropped columns that we didn't need specifically for our ggpairs plot.

```{r Read_Data}
#read in using here pertinent csv file
flower <- read_csv(here("data", "prob2", "shrubstudy_seed_ctwt.ms.data.csv"))
#for gg_miss_var
flower_viz <-flower
flower<-flower%>%
  #drop rows with missing data
  drop_na(nr_seeds)%>%
  clean_names()
flower_subset <- flower %>% 
#create a new dataset of flower, which drops the missing data from nr_seeds because its pertinent to our study
  drop_na(nr_seeds)%>%
  #removed a few columns that were not relevant to ggpairs
  clean_names()%>%
  mutate(date_collected = NULL, shrub_num = NULL, aspect= NULL, tag_num= NULL, plant_nr = NULL, dist_fr_shr = NULL, nr_infl_coll = NULL, wt_seeds_mg = NULL, notes = NULL)

```

```{r exploring_data}
ggplot(data= flower, aes(x=nr_seeds))+
  geom_histogram(bins=13)
#not normal
```



Before cleaning the data, there was missing data for the columns "dist_fr_shr", "wt_seeds_mg", "nr_seeds", and "nr_infil_coll". However, after the data was fully cleaned and wrangled, there was no longer missing data because we dropped the "na" values and removed the observations that weren't neccessary to our statistical analysis. It is important to note that there was missing data from "nr_seeds" and "nr_infil_coll" because these observations are included in our statistical analysis, meaning that these missing values have the potential to skew our results. 

```{r missing_data_viz}
gg_miss_var(flower_viz)

```


```{r skim-dataset}
skim(flower_subset)
```


In order to visualize the relationship between the variables
the ggpairs function was applied. The ggpairs function builds a
scatterplot matrix that displays scatterplots of each pair of numeric
values, while also displaying pearson's r on the right and the variable
distribution on the diagonal.

```{r pairs-plot}
#| fig.width: 12
#| fig.height: 12
#| out.width: 90%
#| fig.align: center
#flower_subset %>%
  #take out site
# select(!species)%>%
ggpairs(flower_subset)
```


We ran multiple regression analysis test to determine the best way to estimate the relationship between the predictor variables, treatment (open or shrub), species, and total number of inflorescence, and the response variable, total number of seeds. We ran linear model, a generalized linear model with a Poisson distribution, generalized linear model with a negative binomial distribution, a generalized linear model with a Poisson distribution and random effect of site, and a generalized linear model with a negative binomial distribution and random effect of site. To check the assumptions of our models, we ran simulateResidual() function in order to compare simulated residuals to observed residuals. We found that flowermodel3 simulated residuals to be the most normal out of all the models. This indicated that indicating that flowermodel3 was likely thre best model to continue moving forward with.

```{r models}
# linear model, we know this is wrong
flowermod1 <- lm(nr_seeds~treatment+species+total_nr_infl, data = flower)

# generalized linear model with Poisson distribution
flowermod2 <- glm(nr_seeds~treatment+species+total_nr_infl, data = flower, family="poisson")

# generalized linear model with negative binomial distribution
flowermod3 <- glm.nb(nr_seeds~treatment+species+total_nr_infl, data = flower)

# generalized linear model with Poisson distribution and random effect of site
flowermod4 <- glmer(nr_seeds~treatment+species+total_nr_infl + (1|shrub_num), data = flower, family = "poisson")

# generalized linear model with negative binomial distribution and random effect of site
flowermod5 <- glmer.nb(nr_seeds~treatment+species+total_nr_infl + (1|shrub_num), data = flower)

```

```{r model-residuals-assumptions}
#| fig.height: 6
#| fig.width: 8
#| fig.align: center

# check diagnostics
simulateResiduals(flowermod1, plot=TRUE)
simulateResiduals(flowermod2, plot=TRUE)
simulateResiduals(flowermod2.a, plot =TRUE)
simulateResiduals(flowermod3, plot=TRUE)
simulateResiduals(flowermod4, plot=TRUE)
simulateResiduals(flowermod5, plot=TRUE)
```


```{r choosing-model}
model.sel(flowermod1,flowermod2, flowermod3, flowermod4, flowermod5)
```

Then we compared the generalized linear model with a negative binomial distribution to a null model to determine if species, and total number of inflorescence impacts the total seed count. The null model is representative of the null hypothesis--the predictor variables, treatment (open or shrub), species, and total number of inflorescence, do not impact the total seed count as it creates a regression model that assumes no relationship between predictor and response variables. On the other hand, the full model creates the negative binomial regression model that assumes a relationship between the predictor variables and the response variable. To check the assumptions of our flowermod3 and null model, we ran simulateResidual() function in order to compare simulated residuals to observed residuals. We found that flowermodel3 still simulated residuals to be the most normal out of all the models. 

```{r null-model-and-assumptions}

null_flower <- glm.nb(nr_seeds~1, data = flower)

simulateResiduals(null_flower, plot= TRUE)

```

```{r AICc-values-for-null-and-mod3}
model.sel(flowermod3, null_flower)
```

Next we ran a few more assumption check for flowermod3 including a test for multicollinearity. We evaluated the multicollinearity by calculating the generalized the variance inflation factor and determined that this model didn't display any multicollnearity because no values we're larger than five. 

```{r Multicollinearity-test}
car::vif(flowermod3)
```  


```{r summaries}
# model object
flowermod3

# summary 
summary(flowermod3)

# confidence intervals
confint(flowermod3)

# adjusted R2
r.squaredGLMM(flowermod3)

```


```{r summary results}
# model object in table
flowermod3%>%
  as_flextable()
```

After picking model3 as the best model, we tested the generalized linear model with a negative binomial with each individual predictor model to test for simpler models. To check the assumptions of our flowermod3 and simpler models, we ran simulateResidual() function in order to compare simulated residuals to observed residuals. We found that flowermodel3 still simulated residuals to be the most normal out of all the models. 

```{r create-simple-models}
flowermod_treatment <- glm.nb(nr_seeds~treatment, data = flower)
flowermod_species <- glm.nb(nr_seeds~species, data = flower)
flowermod_infl <- glm.nb(nr_seeds~total_nr_infl, data = flower)

simulateResiduals(flowermod_treatment, plot= TRUE)
simulateResiduals(flowermod_species, plot= TRUE)
simulateResiduals(flowermod_infl, plot= TRUE)

```

```{r AICc-model_mod3_and_simple}

model.sel(flowermod3, flowermod_infl, flowermod_species, flowermod_treatment)
```



```{r IRR-values}
#getting IRR values using the exp(coef()) function
exp_coef2 <- exp(coef(flowermod3))
exp_coef2

tidy(flowermod3, conf.int = TRUE, conf.level = .95) 
```

## Results 

We found that the flowermod3 including species, treatment, and total number of inflorescence predictors best predicted the total number of seeds because the flowermod3 yields the lowest AIC value indicating that this model has the best predicitive capacity while remaining as simple as possible. With all else held constant, the shrub treatment associated with a 34.7% decrease in the expected total seed count compared to the control treatment (treatmentshrub: IRR= -0.347(95% CI: -0.626,-0.0669) p=0.0161). The Tridas and the Carrup species were found to signifcantly predict the total seed count (p= 0.0423, p<0.001, respectively). On the other hand, the some of the predictors such as, the Gueros, Kobmyo, and Minobt species did not significantly predict total seed count (p= 0.4794, p= 0.3751, p= 0.1272). 

```{r first-vis-maybe}
ggplot(data=flowermod3, aes(x=treatment, y=nr_seeds, fill=predicted))+
  geom_boxplot(position = position_dodge(width=.8), color ="black")+
  labs(x="Species", y =" Seed Count", title = "Distribution of seed count by species and treatment")+
  theme_minimal()+
  facet_wrap(~species)
```


```{r}
flower$residuals <- resid(flowermod3)

# Create a scatter plot of residuals
ggplot(flower, aes(x = predicted, y = residuals)) +
  geom_point(color = "blue", alpha = 0.5, size =2.5) +
  #dashed horizontal line at y=0
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  #adds a linear regression line (without confidence interval) to points, thin to match hline
    geom_smooth(method = "lm", se = FALSE, color = "black", linetype = "dashed", size = .6) +
  #adds a title to graph and to the x and y axis
  labs(x = "Predicted Values", y = "Residuals Values", title = "Residual Analysis") +
  theme_minimal()+
  #Chart title is in bold and size 16
  theme(plot.title = element_text(size = 16, face = "bold"),
          #Axis titles are size 15
        axis.title = element_text(size = 14),
          #Axis text is size 12
        axis.text = element_text(size = 12)
        )

```

Caption: In the graph above, the blue points represent the residual values, the dashed red line represents a reference for if the residual values were equal to '0', and the dashed black line shows a fitted reggression line for the plotted residuals.

Analysis: The negative mean residual value (-.3) illustrates a systematic bias in flowermod3's predictions, which are on average higher than the observed values, indicating that the model is underestimating seed counts by about '.3' seeds. This is also confirmed visually, with a negatively trending regression line below the reference line. Additionally, residuals are not evenly distributed around the reference line, and in conjunction with the BP test illustrates that the assumption of homoscedasticity has been violated.

```{r}
residuals <- resid(flowermod3)
mean_residual <- mean(residuals)
sd_residual <- sd(residuals)

print(mean_residual)
print(sd_residual)

shapiro.test(residuals)
# Breusch-Pagan test for heteroscedasticity
bptest(flowermod3)

```

The p-value (0.002064) illustrates the likelihood of getting the observed test statistic (or a more extreme value) under the assumption of normality. Since the p-value is less than the significance level (0.05), it illustrates that the assumption of normality is violated. Additionally, the tiny p-value for the Breusch-Pagan test (0.00000002694) indicates that the assumption of homoscedasticity is also violated. This highlights that the model underestimates the total seed count with the predictor variables---treatment, total inflourescence, and species---by 0.3 (mean-residual). However, given all of the other tests we ran, we can still consider the generalized linear model with a negative binomial regression was still the most statisically correct is predicting a relationship between the predictor variables and total seed count. More analysis should be done to conclude more statistically significant results. 


## Bibliography

Adams, V.M., Marsh, D.M., & Knox, J.S. (2005). Importance of the seed bank for population viability and population monitoring in a threatened wetland herb, *Biological Conservation*, Volume 124(3), 2005, 425-436. <https://doi.org/10.1016/j.biocon.2005.02.001>

Bakker, J.P., et al. (1996). "Seed Banks and Seed Dispersal: Important Topics in Restoration Ecology." *Acta Botanica Neerlandica*, 45(4), 461-490. <https://doi.org/10.1111/j.1438-8677.1996.tb00806.x>

Giladi, I., Segoli, M., & Ungar, E. D. (2013). Shrubs and herbaceous seed flow in a semi-arid landscape: dual functioning of shrubs as trap and barrier. *Journal of Ecology*, *101*(1), 97--106. <http://www.jstor.org/stable/23354670>

Koontz, T.L., Simpson, H.L. (2010). The composition of seed banks on kangaroo rat (Dipodomys spectabilis) mounds in a Chihuahuan Desert grassland, *Journal of Arid Environments*, 74(10), 2010, 1156-1161. <https://doi.org/10.1016/j.jaridenv.2010.03.008>

Li, J., Li, Y., Qiao, J., Li, L., Wang, X., Yao, J., & Liao, G. (2023). Automatic counting of rapeseed inflorescences using deep learning method and UAV RGB imagery. *Frontiers in Plant Science*, *14*. [https://doi.org/10.3389/fpls.2023.1101143 ](https://doi.org/10.3389/fpls.2023.1101143)

Seaver, M. (2022). *Exploring the Effects of Shrubification on the Community-Level Flowering Phenology and Reproductive Success of Alpine Plants* (Order No. 29069468). Available from ProQuest Dissertations & Theses A&I. (2681729893). <https://www.proquest.com/dissertations-theses/exploring-effects-shrubification-on-community/docview/2681729893/se-2>

Zehnder, T., Lüscher, A., Ritzmann, C. *et al.* (2020). Dominant shrub species are a strong predictor of plant species diversity along subalpine pasture-shrub transects. *Alp Botany*, 130, 141--156. <https://doi.org/10.1007/s00035-020-00241-8>
